{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-11-18 sftrends_manu.Rnw
tables updated with kendall results

A total of twenty-seven \ac{wrtds} models were created, one for each nitrogen at each station.  The functional form of \ac{wrtds} is a simple regression that models the log-transformed response variable as a function of time, flow, and season:

@@

% overall trends
<<trnds>>=
data(trnds_chg)
data(trnds_ave)

# percent changes
tab_chg <- trnds_chg

# increase in bold italic
tab_chg[, -c(1:2)] <- round(tab_chg[, -c(1:2)], 1)
tab_chg[, -c(1:2)][sign(tab_chg[, -c(1:2)]) != -1] <- paste0('\\textit{\\textbf{', tab_chg[, -c(1:2)][sign(tab_chg[, -c(1:2)]) != -1], '}}')
tab_chg <- gather(tab_chg, 'var', 'chg', -Site_Code, -resvar) %>% 
  mutate(chg = paste0('\\footnotesize{(', chg, ')}'))

# averages
tab_ave <- gather(trnds_ave, 'var', 'ave', -Site_Code, -resvar) %>% 
  mutate(ave = round(ave, 1))

tab <- left_join(tab_ave, tab_chg, by = c('Site_Code', 'resvar', 'var')) %>% 
  unite(val, ave, chg, sep = ' ') %>% 
  spread(var, val) %>% 
  arrange(resvar, Site_Code) %>% 
  select(Site_Code, resvar, `1976-1995`, `1996-2014`, Spring, Summer, Fall, Winter)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are averages (mg L$^{-1}$) and percent changes in parentheses (increasing in bold-italic).  Changes are based on the difference between the ending and starting estimates within each period.  See \\cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual', 'Seasonal'),
  n.cgroup = c(2, 4),
  rowname = stats,
  label = 'tab:trnds'
  )

@@

data(h2dat)
dt <- as.Date('2007-05-01')

@@

mobrks <- list(c(3, 4, 5), c(6, 7, 8), c(9, 10, 11), c(12, 1, 2))
yrbrks_bf <- c(1975, 2007)
yrbrks_af <- c(2006, 2013)
molabs <- c('Spring', 'Summer', 'Fall', 'Winter')
yrlabs_bf <- '1976-2007'
yrlabs_af <- '2008-2013'
faclevs <- c('1976-2007_before', '2008-2013_after', paste0(molabs, '_before'), paste0(molabs, '_after'))

# summarize din, sio2, chl trends
no_bf <- wrtdstrnd(noraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'before', var = 'no')
no_af <- wrtdstrnd(noraw_af, mobrks, yrbrks_af, molabs, yrlabs_af, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'after', var = 'no')
nh_bf <- wrtdstrnd(nhraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'before', var = 'nh')
nh_af <- wrtdstrnd(nhraw_af, mobrks, yrbrks_af, molabs, yrlabs_af, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'after', var = 'nh')

# combine for trends for table
totab <- rbind(no_bf, no_af, nh_bf, nh_af) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 2))
    ) %>% 
  gather('est', 'val', chg:ave) %>% 
  unite('col', var, est) %>% 
  spread(col, val) %>% 
  unite('period', cat, period) %>% 
  mutate(period = factor(period, levels = faclevs, labels = faclevs)) %>% 
  arrange(period) %>% 
  mutate(cat = c(yrlabs_bf, yrlabs_af, molabs, molabs)) %>% 
  data.frame(sep = c(rep('Annual', 2), rep('Seasonal, pre', 4), rep('Seasonal, post', 4)), .) %>% 
  select(sep, cat, no_ave, no_chg, nh_ave, nh_chg)

# final table formatting
segs <- unique(totab$sep)
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('NO$_{2}^{-}$/NO$_{3}^{2-}$', 'NH$_{4}^{+}$')
names(totab) <- rep(c('Mean', '\\% change'), 2)

cap.val<-'Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \\ac{wwtp} upgrades upstream of station P8. Upgrades were completed in 2006 at the City of Stockton \\ac{wwtp} (San Joaquin County, \\cref{fig:stock}).  Summaries are means and percent changes based on annual means within the pre- and post-upgrade time periods (1976-2007, 2008-2013).  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(2, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:p8chg'
  )

@@

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 1))
    ) %>% 
  gather('var', 'val', chg:ave) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_ave, din_chg, chl_ave, chl_chg, sio_ave, sio_chg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Mean', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:d7chg'
  )

@@

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 1))
    ) %>% 
  gather('var', 'val', chg:ave) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_ave, din_chg, chl_ave, chl_chg, sio_ave, sio_chg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Mean', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:c10chg'
  )

>>>>

A total of twenty-seven \ac{wrtds} models were created, one for each nitrogen analyte at each station.  The functional form of \ac{wrtds} is a simple regression that models the log-transformed response variable as a function of time, flow, and season:

@@

% overall trends
<<trnds>>=
data(trnds_sk)

# percent changes
tab <- trnds_sk %>% 
  select(Site_Code, resvar, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      paste0('(', perchg, ')' ), 
      paste0('(\\textit{\\textbf{', perchg, '}})')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  mutate(
    perchg = paste0('\\footnotesize{', perchg, '}')
  ) %>% 
  unite('val', med, perchg, sep = ' ') %>% 
  spread(cat, val) %>% 
  arrange(resvar, Site_Code) %>% 
  select(Site_Code, resvar, `1976-1995`, `1996-2014`, Spring, Summer, Fall, Winter)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold-italic). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. See \\cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual', 'Seasonal'),
  n.cgroup = c(2, 4),
  rowname = stats,
  label = 'tab:trnds',
  insert.bottom = foot.val
  )

@@

data(h2dat)
dt <- as.Date('2006-01-01')

@@

mobrks <- list(c(3, 4, 5), c(6, 7, 8), c(9, 10, 11), c(12, 1, 2))
yrbrks_bf <- c(1975, 2005)
yrbrks_af <- c(2006, 2013)
molabs <- c('Spring', 'Summer', 'Fall', 'Winter')
yrlabs_bf <- '1976-2006'
yrlabs_af <- '2007-2013'
faclevs <- c('1976-2006_before', '2007-2013_after', paste0(molabs, '_before'), paste0(molabs, '_after'))

# summarize din, sio2, chl trends
no_bf <- wrtdstrnd_sk(noraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf) %>% 
  mutate(period = 'before', var = 'no')
no_af <- wrtdstrnd_sk(noraw_af, mobrks, yrbrks_af, molabs, yrlabs_af) %>% 
  mutate(period = 'after', var = 'no')
nh_bf <- wrtdstrnd_sk(nhraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf) %>% 
  mutate(period = 'before', var = 'nh')
nh_af <- wrtdstrnd_sk(nhraw_af, mobrks, yrbrks_af, molabs, yrlabs_af) %>% 
  mutate(period = 'after', var = 'nh')

# combine for trends for table
totab <- rbind(no_bf, no_af, nh_bf, nh_af) %>% 
  select(-tau, -slope, -chitest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      perchg, 
      paste0('\\textit{\\textbf{', perchg, '}}')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  gather('est', 'val', med:perchg) %>% 
  unite('col', var, est) %>% 
  spread(col, val) %>% 
  unite('period', cat, period) %>% 
  mutate(period = factor(period, levels = faclevs, labels = faclevs)) %>% 
  arrange(period) %>% 
  mutate(cat = c(yrlabs_bf, yrlabs_af, molabs, molabs)) %>% 
  data.frame(sep = c(rep('Annual', 2), rep('Seasonal, pre', 4), rep('Seasonal, post', 4)), .) %>% 
  select(sep, cat, no_med, no_perchg, nh_med, nh_perchg)

# final table formatting
segs <- unique(totab$sep)
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('NO$_{2}^{-}$/NO$_{3}^{2-}$', 'NH$_{4}^{+}$')
names(totab) <- rep(c('Median', '\\% change'), 2)

cap.val<-'Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \\ac{wwtp} upgrades upstream of station P8. Upgrades were completed in 2006 at the City of Stockton \\ac{wwtp} (San Joaquin County, \\cref{fig:stock}).  Summaries are  medians and percent change per year in parentheses (increasing in bold-italic).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(2, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:p8chg',
  insert.bottom = foot.val
  )

@@

# summarize din, sio2, chl trends
din <- wrtdstrnd_sk(dinraw, mobrks, yrbrks, molabs, yrlabs)
din <- wrtdstrnd_sk(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd_sk(chlraw, mobrks, yrbrks, molabs, yrlabs)
chl <- wrtdstrnd_sk(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd_sk(sioraw, mobrks, yrbrks, molabs, yrlabs)
sio <- wrtdstrnd_sk(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  select(res, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      perchg, 
      paste0('\\textit{\\textbf{', perchg, '}}')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  gather('var', 'val', med:perchg) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_med, din_perchg, chl_med, chl_perchg, sio_med, sio_perchg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Median', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are  medians and percent change per year in parentheses (increasing in bold-italic).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:d7chg', 
  insert.bottom = foot.val
  )

@@

# summarize din, sio2, chl trends
din <- wrtdstrnd_sk(dinraw, mobrks, yrbrks, molabs, yrlabs)
din <- wrtdstrnd_sk(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd_sk(chlraw, mobrks, yrbrks, molabs, yrlabs)
chl <- wrtdstrnd_sk(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd_sk(sioraw, mobrks, yrbrks, molabs, yrlabs)
sio <- wrtdstrnd_sk(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013')) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

# combine for trends for table
totab <- rbind(din, chl, sio) %>% 
  select(res, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      perchg, 
      paste0('\\textit{\\textbf{', perchg, '}}')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  gather('var', 'val', med:perchg) %>% 
  unite('col', res, var) %>% 
  spread(col, val) %>% 
  mutate(cat = factor(cat, levels = c('1976-2013', yrlabs, molabs), labels = c('1976-2013', yrlabs, molabs))) %>% 
  arrange(cat) %>% 
  data.frame(sep = c(rep('Annual', 5), rep('Seasonal', 4)), .) %>% 
  select(sep, cat, din_med, din_perchg, chl_med, chl_perchg, sio_med, sio_perchg)

# final table formatting
segs <- c('All', 'Annual', 'Seasonal')
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('DIN', 'Chl-\\textit{a}', 'SiO$_2$')
names(totab) <- rep(c('Median', '\\% change'), 3)

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station C10. Summaries are  medians and percent change per year in parentheses (increasing in bold-italic).  Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. Increasing values are in bold-italics. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(1, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2, 2),
  rowname = cats,
  label = 'tab:c10chg', 
  insert.bottom = foot.val
  )
